---
title: "Step2_Data_Integration_And_Clustering"
output: html_document
---
```{r,setup, results='hide'}
seed=1234
finalfilt_noribo_noemptywells<- readRDS("finalfilt_noribo_may2022.rds")
dim(finalfilt_noribo_noemptywells)
```

#Normalization, Scaling and PCA Analysis
```{r,message='hide',warning='hide', results='hold',results='hold',fig.height=5,fig.width=16}
finalfilt_noribo_noemptywells = NormalizeData(finalfilt_noribo_noemptywells) #by default uses scale.factor=10000
finalfilt_noribo_noemptywells_genes <- rownames(finalfilt_noribo_noemptywells)
finalfilt_noribo_noemptywells <- ScaleData(finalfilt_noribo_noemptywells, features =finalfilt_noribo_noemptywells_genes)

suppressWarnings(suppressMessages(finalfilt_noribo_noemptywells <- FindVariableFeatures(finalfilt_noribo_noemptywells, selection.method = "vst", nfeatures = 2000 ,verbose = FALSE,assay = "RNA")))
top20 <- head(VariableFeatures(finalfilt_noribo_noemptywells), 20)

LabelPoints(plot = VariableFeaturePlot(finalfilt_noribo_noemptywells), points = top20, repel = TRUE)
finalfilt_noribo_noemptywells <- RunPCA(finalfilt_noribo_noemptywells, npcs = 50, verbose = F) #by default uses the Variable features


library(cowplot)
plot_grid(ncol = 3,
  DimPlot(finalfilt_noribo_noemptywells, reduction = "pca", group.by = "orig.ident",dims = 1:2),
  DimPlot(finalfilt_noribo_noemptywells, reduction = "pca", group.by = "orig.ident",dims = 3:4),
  DimPlot(finalfilt_noribo_noemptywells, reduction = "pca", group.by = "orig.ident",dims = 5:6) )
VizDimLoadings(finalfilt_noribo_noemptywells, dims = 1:5, reduction = "pca",ncol = 5,balanced = T)
ElbowPlot(finalfilt_noribo_noemptywells, reduction = "pca",ndims = 50)
finalfilt_noribo_noemptywells <- RunUMAP(finalfilt_noribo_noemptywells, reduction = "pca", dims = 1:20,
                   n.components=2,
                   n.neighbors=30,
                   n.epochs=200,
                   min.dist=0.3,
                   learning.rate=1,
                   spread=1 )
DimPlot(finalfilt_noribo_noemptywells,group.by = "seq_batch") # here we see there is a lot of batch effect in our data

```

#Data integration using Harmony Package
We integrate the data due to batch effects since patients were sequenced in different batches and we want to get rid of this and keep only biological contrubution
Dimensions to be used fro chosen 20 from the Elbow Plot
```{r,message='hide',warning='hide', results='hold',results='hold',fig.height=5,fig.width=16}
library(harmony)

finalfilt_noribo_noemptywells_harmony <- RunHarmony(
  finalfilt_noribo_noemptywells,
  group.by.vars = "seq_batch",
  dims.use = 1:20,
  plot_convergence = TRUE,max.iter.harmony = 8) # number of iterations were chosen based on the convergence plot, after 8 there isnt any more contribution so 8 was enough

finalfilt_noribo_noemptywells_harmony<-FindNeighbors(finalfilt_noribo_noemptywells_harmony,dims=1:20, reduction = "harmony",assay = "RNA")
finalfilt_noribo_noemptywells_harmony <- RunUMAP(finalfilt_noribo_noemptywells_harmony, dims = 1:20, reduction = "harmony", reduction.name = "umap_harmony",return.model = T)
#supplemantary figure for before& after data integration
p1 <- DimPlot(finalfilt_noribo_noemptywells_harmony, reduction = "umap", group.by = "seq_batch")+ggtitle("UMAP raw_data")
p2 <- DimPlot(finalfilt_noribo_noemptywells_harmony, reduction = "umap_harmony", group.by = "seq_batch")+ggtitle("UMAP Harmony")
leg <- get_legend(p1)

gridExtra::grid.arrange(
  gridExtra::arrangeGrob(
    p1 + NoLegend() + NoAxes(),
    p2 + NoLegend() + NoAxes(),ncol = 2),
  leg, ncol=2,widths=c(8,2)
)

p1 <- DimPlot(finalfilt_noribo_noemptywells_harmony, reduction = "umap_harmony", group.by = "seq_batch")+ggtitle("harmony")
p2 <- DimPlot(finalfilt_noribo_noemptywells_harmony, reduction = "umap", group.by = "seq_batch")+ggtitle("no harmony")
p1+p2

#saveRDS(finalfilt_noribo_noemptywells_harmony,"May2022scewithtreated.rds")
```

#Clustering 
We remove the cells from the second biopsies and only keep "naive" cells
Afterwards we look at differet resolutions for the louvain algorithm.We looked at the difference of clustering in different resolutions and decided 1.5 was the best with biologically meaningful clusters.
```{r,message='hide',warning='hide', results='hold',results='hold',fig.height=5,fig.width=16}
finalfilt_noribo_noemptywells_harmony.naive<-subset(finalfilt_noribo_noemptywells_harmony,subset=treatment.status=="naive")
DimPlot(finalfilt_noribo_noemptywells_harmony.naive, reduction = "umap_harmony", group.by = "seq_batch")


finalfilt_noribo_noemptywells_harmony.naive<-FindNeighbors(finalfilt_noribo_noemptywells_harmony.naive,dims=1:20, reduction = "harmony",assay = "RNA")

#finalfilt_noribo.harmony.naive <- RunUMAP(finalfilt_noribo.harmony.naive, dims = 1:20, reduction = "harmony", reduction.name = "umap_harmony",return.model = T)
library(pheatmap)
pheatmap(finalfilt_noribo_noemptywells_harmony.naive@graphs$RNA_snn[1:200,1:200],
         col=c("white","black"),border_color = "grey90",
         legend = F,cluster_rows = F,cluster_cols = F,fontsize = 2)

# Clustering with louvain (algorithm 1)
for (res in c( 0.1 , 0.25 , .5 , 1 , 1.5 , 2 )){
  finalfilt_noribo_noemptywells_harmony.naive <- FindClusters(finalfilt_noribo_noemptywells_harmony.naive, graph.name = "RNA_snn", resolution = res , algorithm = 1)
}

plot_grid(ncol = 3,
          DimPlot(finalfilt_noribo_noemptywells_harmony.naive, reduction = "umap_harmony", group.by= "RNA_snn_res.1")+ggtitle("Harmony_louvain_1"),
                  DimPlot(finalfilt_noribo_noemptywells_harmony.naive, reduction = "umap_harmony", group.by = "RNA_snn_res.1.5")+ggtitle("Harmony_louvain_1.5"),
                  DimPlot(finalfilt_noribo_noemptywells_harmony.naive, reduction = "umap_harmony", group.by = "RNA_snn_res.2")+ggtitle("Harmony_louvain_2")
          )

clustree(finalfilt_noribo_noemptywells_harmony.naive)

finalfilt_noribo_noemptywells_harmony.naive.resol1.5<-FindClusters(finalfilt_noribo_noemptywells_harmony.naive, graph.name = "RNA_snn", resolution = 1.5 , algorithm = 1)
DimPlot(finalfilt_noribo_noemptywells_harmony.naive.resol1.5,reduction = "umap_harmony")
May2022sce<-finalfilt_noribo_noemptywells_harmony.naive.resol1.5
```


#The Colors we have chosen for the clusters and tissues
```{r,fig.height=5,fig.width=10}
library(RColorBrewer)
library(scales)
safe_colorblind_palette <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
                             "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888")
show_col(safe_colorblind_palette)

pal <- c("#000000","#004949","#009292","#ff6db6","#ffb6db",
 "#490092","#006ddb","#b66dff","#6db6ff","#b6dbff",
 "#920000","#924900","#db6d00","#24ff24","#ffff6d")
show_col(pal)

#darkolivegreen1

myositiscolors<-c("#999933","#004949","#009292","#ff6db6","#ffb6db",
 "#490092","#006ddb","#b66dff","#6db6ff","#b6dbff",
 "#920000","#924900","#db6d00","darkolivegreen1","lightsalmon","burlywood2")
show_col(myositiscolors)
#https://www.colorcodepicker.com/
myositiscolorsforditto<-c("#f4847c","#0bbec5","#999933","#004949","#009292","#ff6db6","#ffb6db",
 "#490092","#006ddb","#b66dff","#6db6ff","#b6dbff",
 "#920000","#924900","#db6d00","darkolivegreen1","lightsalmon","burlywood2")

my_cols <- c("0"="#999933","1"="#004949","2"="#009292","3"="#ff6db6","4"="#ffb6db",
 "5"="#490092","6"="#006ddb","7"="#b66dff","8"="#6db6ff","9"="#b6dbff",
 "10"="#920000","11"="#924900","12"="#db6d00","13"="burlywood2")

my_cols2<- my_cols[order(as.integer(names(my_cols)))]



colrs <- brewer.pal.info[brewer.pal.info$colorblind == TRUE, ]
col_vec = unlist(mapply(brewer.pal, colrs$maxcolors, rownames(colrs)))
show_col(col_vec)
DimPlot(finalfilt_noribo_noemptywells_harmony.naive.resol1.5, reduction = "umap_harmony",cols=my_cols2)

tissue<-c("MUSL"="royalblue3","PB"="seagreen3")
DimPlot(finalfilt_noribo_noemptywells_harmony.naive.resol1.5, reduction = "umap_harmony",group.by = "tissue", cols = tissue)

#332288
#AA4499
```


#Finding DE genes in clusters
```{r,message='hide',warning='hide', results='hold',results='hold',fig.height=5,fig.width=16}

#we preivously selected resol1.5 for the clustering so we only use that one

markers_genes_resol1.5 <- FindAllMarkers(May2022sce,
                               log2FC.threshold = 0.2,
                               test.use = "wilcox",
                               min.pct = 0.1,
                               min.diff.pct = 0.2,
                               only.pos = TRUE,
                               max.cells.per.ident = 50,
                               assay = "RNA") # all the parameters are default

markers_genes_resol1.5 %>% group_by(cluster)  %>% top_n(-5, p_val) -> top10_resol1.5
top10_resol1.5
top5genes<-top10_resol1.5$gene
top5genes<-c(top5genes,"CD4","CD8B","CD8A") # We also add amnually the CD4, CD8 genes to the heatmap

mypar(2,5,mar=c(4,6,3,1))
for(i in unique(top10_resol1.5$cluster)){
  barplot( sort( setNames(top10_resol1.5$avg_log2FC, top10_resol1.5$gene) [top10_resol1.5$cluster == i], F),
           horiz = T,las=1 ,main=paste0(i," vs. rest"),border = "white", yaxs="i" )
  abline(v=c(0,0.25),lty=c(1,2))
}

finalfilt_noribo_noemptywells_harmony.naive.resol1.5.copy <- ScaleData(finalfilt_noribo_noemptywells_harmony.naive.resol1.5, features = unique(top5genes), assay = "RNA") # the data is scaled based on the top5genes for visualization purposes


DoHeatmap(finalfilt_noribo_noemptywells_harmony.naive.resol1.5.copy, features = unique(top5genes),label=TRUE,group.colors =my_cols2)+scale_fill_viridis()+ theme(text = element_text(size = 10))
#based on clusters and DE genes we go with resolution 1.5


#dittoHeatmap(finalfilt_noribo_noemptywells_harmony.naive.resol1.5.copy,unique(top10_resol1.5$gene), annot.by=c("tissue","seurat_clusters"),order.by=c("tissue","seurat_clusters"),annot.colors = myositiscolorsforditto,scaled.to.max = TRUE, heatmap.colors.max.scaled = colorRampPalette(c("darkblue","mediumaquamarine" ,"yellow"))(25))

#write.csv(top10_resol1.5,"top10_resol1.5.csv")
#write.csv(markers_genes_resol1.5,"markers_genes_resol1.5.csv")
#saveRDS(May2022sce,"May2022sce.rds")
```

#DE genes between Muscle vs Blood cells
```{r,fig.height=5,fig.width=10}
# Compare muscle vs blood
May2022sce.copy<-May2022sce
tissue_comparision <- SetIdent(May2022sce.copy, value = "tissue") # first we set the idents to be tissues instead of clusters

#Compute differential expression
genespertissue<-FindMarkers(tissue_comparision, ident.1="MUSL", ident.2="PB",min.pct=0.2) # min.pct to match the DE in FindAllMarkers
genespertissue$genes<-rownames(genespertissue)
genespertissue   %>% top_n(-50, p_val) -> top50_genes
genespertissue   %>% top_n(-5, p_val) -> top5_genes

top5_genes$genes<-rownames(top5_genes)
top50_genes$genes<-rownames(top50_genes)
top50_gene_names<-rownames(top50_genes)
top50_gene_names<-c(top50_gene_names,"CD4","CD8B")
May2022sce.copy2 <- ScaleData(May2022sce.copy, features = as.character(unique(top50_gene_names)), assay = "RNA")
DoHeatmap(May2022sce.copy2, features = unique(top50_genes$genes),group.by = "tissue")+scale_fill_viridis()
DoHeatmap(May2022sce.copy2, features = unique(top50_gene_names),group.by = "tissue")+scale_fill_gradient2( low = rev(c('#d1e5f0','#67a9cf','#2166ac')), mid = "white", high = rev(c('#b2182b','#ef8a62','#fddbc7')), midpoint = 0, guide = "colourbar", aesthetics = "fill")



#write.csv(genespertissue,"genespertissue_210422.csv")
VlnPlot(May2022sce.copy2, features = c("CXCR4","LMNA","CREM","TXNIP"),ncol = 2,group.by = "tissue",assay = "RNA", pt.size = .1,cols = tissue)

FeaturePlot(May2022sce,features = c("CD3D"),reduction="umap_harmony",cols = c("grey", "dodgerblue4"))
FeaturePlot(May2022sce,features = c("CD4"),reduction="umap_harmony",cols=c("grey","dodgerblue4"))
FeaturePlot(May2022sce,features = c("CD8B"),reduction="umap_harmony")
FeaturePlot(May2022sce,features = c("CD8A"),reduction="umap_harmony")
FeaturePlot(May2022sce,features = c("CD8B"),reduction="umap_harmony")
plot_density(May2022sce, "CD4",reduction = "umap_harmony")
plot_density(May2022sce, "CD3D",reduction = "umap_harmony")
plot_density(May2022sce, "CD8A",reduction = "umap_harmony")
plot_density(May2022sce, "CD8B",reduction = "umap_harmony")


```
